# Course Project 1

## Read in the data

```{r readdata, echo=TRUE}
library(dplyr)
library(ggplot2)
step_data<-read.csv("activity.csv")
```
## What is the mean total number of steps taken per day

Histogram of total number of steps taken each day

```{r hist, echo=TRUE}
grouped_data<-step_data%>%group_by(date)%>%summarise(tot_steps=sum(steps,na.rm=TRUE))
hist(grouped_data$tot_steps,xlab="Total Steps by Day",main="Histogram of Total Steps by Day")
mean_steps<-format(mean(grouped_data$tot_steps),scientific=FALSE)
median_steps<-format(median(grouped_data$tot_steps),sceintific=FALSE)
```

The mean total number of steps per day is `r mean_steps`. The median number of steps per day is `r median_steps`.  

## What is the average daily pattern?

Time series plot of the 5-minute intervals and the average number of total steps taken

```{r meansteps,echo=TRUE}
grouped_data<-step_data%>%group_by(interval)%>%summarise(avg_steps=mean(steps,na.rm=TRUE))
plot(grouped_data$interval,grouped_data$avg_steps,type="l",xlab="Interval",ylab="Average Steps",main="Average Steps by 5-minute Interval")
max_interval<-grouped_data[which.max(grouped_data$avg_steps),1]
```
The 5-minute interval with the maximum average number of steps is interval `r max_interval`.

## Imputing missing values

```{r countNA,echo=TRUE}
num_NA<-sum(is.na(step_data$steps))
```

There are `r num_NA` missing values in the dataset.

We will impute the missing values using the mean of the interval of the missing value. Here is a histogram of the dataset with imputed values.

```{r impute,echo=TRUE}
imputed_data<-step_data%>%group_by(interval)%>%mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=TRUE),steps))
grouped_imputed_data<-imputed_data%>%group_by(date)%>%summarise(tot_steps=sum(steps))
hist(grouped_imputed_data$tot_steps,xlab="Total Steps by Day",main="Histogram of Total Steps by Day")
mean_imputed_steps<-format(mean(grouped_imputed_data$tot_steps),scientific=FALSE)
median_imputed_steps<-format(median(grouped_imputed_data$tot_steps),scientific=FALSE)
```

After imputing the missing values, the mean total number of steps per day is `r mean_imputed_steps`, and the median number of steps per day is `r median_imputed_steps`.

The impact of imputing values brings the mean and median number of steps per day up. 

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday,echo=TRUE}
imputed_data$date<-as.Date(imputed_data$date)
weekday_data<-imputed_data%>%mutate(wkdy=ifelse(weekdays(date)=="Sunday"|weekdays(date)=="Saturday","Weekend","Weekday"))
#imputed_weekday_data<-weekday_data%>%group_by(interval)%>%mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=TRUE),steps))
grouped_imputed_weekday_data<-weekday_data%>%group_by(interval,wkdy)%>%summarise(avg_steps=mean(steps,na.rm=TRUE))
qplot(interval,avg_steps,data=grouped_imputed_weekday_data,facets=.~wkdy,geom="line",xlab="Interval",ylab="Average number of steps taken")
